\documentclass[12pt]{extarticle}
\usepackage[utf8]{inputenc}
\usepackage{hyperref}
\usepackage{enumitem}
\usepackage{parskip}
\usepackage{setspace}
\usepackage{titlesec}
\usepackage{amsmath}
\usepackage[margin=0.5in]{geometry}
\usepackage{fancyhdr}

\title{\Huge Basic B\&W Film Developing Guide}
\author{\Large Jimmy Ha}
\date{\large 29 September, 2025}

\pagestyle{fancy}
\fancyhf{}
\fancyfoot[C]{\thepage}
\renewcommand{\headrulewidth}{0pt}

\begin{document}

\maketitle

\vspace{-2em}

This guide is an introduction to film developers. A lot of experimentation is possible, and the different results can be very noticeable. When starting out, it’s best to stick with common 400-speed films and common developers, because the more widespread something is, the more information is available. \textbf{If you don't know where to start, then get Ilford HP5+}, and develop with either Kodak D-76, Kodak HC-110 or Sprint Standard.

\begin{table}[h!]
\centering
\begin{tabular}{|l|c|c|c|c|}
\hline
\multicolumn{5}{|c|}{Sprint 1+9} \\
\hline
Film & 20ºC & 21ºC & 22ºC & 23ºC \\
\hline
Fomapan 400 & 13:00 & 11:45 & 10:15 & 9:30 \\
Ilford HP5+ 400 & 10:00 & 9:00 & 8:00 & 7:15 \\
Kentmere 400 & 11:30 & & & \\
Kodak T-Max 400 & 11:30 & 10:15 & 9:15 & 8:30 \\
Kodak Tri-X 400 & 10:00 & 9:00 & 8:00 & 7:15 \\
Kodak TXP 320 & 8:30 & 7:45 & 7:00 & 6:15 \\
\hline
\end{tabular}
\end{table}

\begin{table}[h!]
\centering
\begin{tabular}{|l|c|c|c|c|}
\hline
\multicolumn{5}{|c|}{Kodak D-76 1+1} \\
\hline
Film & 20ºC & 21ºC & 22ºC & 23ºC \\
\hline
Fomapan 400 & 12:00 & & & \\
Ilford HP5+ 400 & 13:00 & & & \\
Kentmere 400 & 14:00 & & & \\
Kodak T-Max 400 & 12:30 & 11:00 & 10:00 & 9:00 \\
Kodak Tri-X 400 & 10:00 & 9:30 & 9:00 & 8:00 \\
Kodak TXP 320 & 9:00 & & & \\
\hline
\end{tabular}
\end{table}

\begin{table}[h!]
\centering
\begin{tabular}{|l|c|c|c|c|}
\hline
\multicolumn{5}{|c|}{Kodak HC-110 1+31 Dilution B} \\
\hline
Film & 20ºC & 21ºC & 22ºC & 23ºC \\
\hline
Fomapan 400 & 7:00 & & & \\
Ilford HP5+ 400 & 5:00 & & & \\
Kentmere 400 & 5:30 & & & \\
Kodak T-Max 400 & 6:00 & 5:30 & 5:00 & 4:30 \\
Kodak Tri-X 400 & 3:45 & 3:30 & 3:00 & 2:30 \\
Kodak TXP 320 & 4:45 & 4:15 & 4:00 & 3:30 \\
\hline
\end{tabular}
\end{table}

\pagebreak

\subsection*{Basic Developing Questions}

\subsubsection*{What does 1+1, 1+31, etc mean?}
\begin{itemize}
\item That value represents dilution. It does not represent ratio.
\item Example: "Sprint 1:9" means you have 50mL concentrate and 450mL water.
\item A high dilution let you save money by using less concentrate; the tradeoff is longer developing times and less contrast.
\end{itemize}

\subsubsection*{What is the correct way to agitate?}
When developing film, follow Ilford's method: 2 inversions every 5 seconds (see \url{https://www.ilfordphoto.com/beginners-guide-processing-film/}).
\begin{itemize}
\item If the instructions say “agitate for 10 seconds,” then do 4 inversions.
\item If the instructions say “agitate for 30 seconds,” then do 12 inversions.
\item After doing inversions, tap the tank against the table to dislodge any air bubbles.
\item Kodak has slightly different rules for agitation.
\end{itemize}

\subsubsection*{When do I start and stop the timer?}
\begin{itemize}
\item Ilford recommends starting the timer as soon as all chemicals are in the tank.
\item Ilford also recommends pouring chemicals out 15 seconds before the timer ends.
\item Do not develop anything under 5 minutes. Adjust developer temperature if necessary.
\end{itemize}

\subsection*{Basic Developing Steps}

Instructions combine Ilford's official method and Bob Laubach's guidance. Consistency is key.

\subsubsection*{1. Prewash}
\begin{itemize}
\item 60 seconds of water inversions
\item Required for 120 film
\item Pour out water into drain
\end{itemize}

\subsubsection*{2. Develop}
\begin{itemize}
\item Pour chemicals into tank and start timer
\item Agitate continuously first 30 seconds (12 inversions)
\item Let rest
\item At start of each minute, agitate 10 seconds (4 inversions)
\item When timer stops, pour chemicals into drain
\end{itemize}

\subsubsection*{3. Stop Bath}
\begin{itemize}
\item Pour stop bath into tank
\item Timer: 30 seconds, agitate continuously
\item Pour down drain
\end{itemize}

\subsubsection*{4. Fixer}
\begin{itemize}
\item Pour fixer into tank
\item Timer: 5 minutes
\item Agitate continuously first 3 minutes (30 inversions per minute)
\item Agitate 10s per minute afterwards
\item Pour into recycling drain, not regular drain
\end{itemize}

\subsubsection*{5. Wash}
\begin{itemize}
\item Open lid, timer 5 minutes
\item Fill tank under running water, let sit a few seconds
\item Repeat
\end{itemize}

\subsubsection*{6. Final Rinse}
\begin{itemize}
\item Optional, recommended
\item Timer: 60 seconds
\item Fill with wetting agent (e.g., Photo Flo)
\item Spin reel a few rotations
\end{itemize}

\subsection*{Film Reviews}

\subsubsection*{ISO 400 films rankings}
\begin{itemize}
\item 1st place = smoothest linear contrast, not highest contrast
\item Prices vary by source
\end{itemize}

\begin{table}[h!]
\centering
\begin{tabular}{|l|l|l|l|l|l|}
\hline
Rank & Dynamic Range & Fine Grain & Contrast & Sharpness & Price \\
\hline
Best & Kodak Tri-X & Kodak T-Max & Kodak T-Max & Kodak T-Max & Foma \\
2nd & Ilford HP5+ & Ilford Delta 400 & Ilford Delta 400 & Ilford Delta 400 & Kentmere \\
3rd & Kodak T-Max & Kodak Tri-X & Ilford HP5+ & Kodak Tri-X & Ilford HP5+ \\
4th & Ilford Delta & Ilford HP5+ & Kodak Tri-X & Ilford HP5+ & Ilford Delta \\
5th & Kentmere & Kentmore & Kentmere & Kentmere & Kodak Tri-X \\
Last & Foma & Foma & Foma & Foma & Kodak T-Max \\
\hline
\end{tabular}
\end{table}

\subsubsection*{ISO 400 Films Summary}
Push/Pull allows using a different ISO for lighting: push = higher ISO (e.g., 800), pull = lower ISO (e.g., 200).

\paragraph{Arista EDU Ultra 400}
\begin{itemize}
\item Not real, relabeled Fomapan 400
\end{itemize}

\paragraph{Foma Fomapan 400}
\begin{itemize}
\item 4x5 available
\item High contrast, narrow dynamic range, limited push/pull, cheap
\item Use only if experienced
\end{itemize}

\paragraph{Ilford Delta 400}
\begin{itemize}
\item 4x5 available
\item Good dynamic range, small grain, decent push/pull, cheaper than T-Max
\end{itemize}

\paragraph{Ilford HP5+ 400}
\begin{itemize}
\item 4x5 available
\item Excellent dynamic range and push/pull
\item Good for beginners
\end{itemize}

\paragraph{Kentmere Pan 400}
\begin{itemize}
\item No 4x5
\item Decent dynamic range and push/pull, flimsy base
\item Cheaper alternative to HP5+
\end{itemize}

\paragraph{Kodak T-Max}
\begin{itemize}
\item 4x5 available
\item Technically high quality: smallest grain, best sharpness, best contrast
\item More expensive than Tri-X
\end{itemize}

\paragraph{Kodak Tri-X}
\begin{itemize}
\item 4x5 available for ISO 320
\item Best dynamic range and push/pull
\item Comparable to HP5+ but more contrast
\end{itemize}

\subsection*{Useful Links}
\begin{itemize}
\item The Massive Dev Chart: \url{https://www.digitaltruth.com/devchart.php}
\end{itemize}

\subsection*{Technical Data Sheets}

\paragraph{Developers}
\begin{itemize}
\item Agfa Rodinal: \url{https://www.digitaltruth.com/products/agfa_tech/Rodinal_Leaflet.pdf}
\item Ilford Ilfotec DD-X: \url{https://www.ilfordphoto.com/wp/wp-content/uploads/2019/08/ILFOTEC-DDX-AUG19.pdf}
\item Ilford Perceptol: \url{https://www.ilfordphoto.com/amfile/file/download/file/1829/product/551/}
\item Kodak D-76: \url{https://business.kodakmoments.com/sites/default/files/files/resources/j78.pdf}
\item Kodak XTOL: \url{https://www.kodakprofessional.com/sites/default/files/wysiwyg/pro/chemistry/J-109_Feb_2018.pdf}
\item Kodak HC-110: \url{https://www.kodakprofessional.com/sites/default/files/wysiwyg/pro/chemistry/j24.pdf}
\item Kodak T-Max 400: \url{https://www.kodakprofessional.com/sites/default/files/wysiwyg/pro/chemistry/j86.pdf}
\item Sprint Standard: \url{https://static1.squarespace.com/static/5e2f0253df34017bdcf28821/t/5eaf5cff2e269141d85ecd9a/1588550913997/Standard_TimeChart.pdf}
\end{itemize}

\paragraph{Films}
\begin{itemize}
\item Foma Fomapan 400: \url{https://www.foma.cz/en/fomapan-400}
\item Ilford tables: \url{https://www.ilfordphoto.com/wp/wp-content/uploads/2017/03/Film-processing-chart-.pdf}
\item Ilford HP5+: \url{https://www.ilfordphoto.com/amfile/file/download/file/1953/product/691/}
\item Kentmere Pan 400: \url{https://www.ilfordphoto.com/amfile/file/download/file/1959/product/2136/}
\item Kodak T-Max: \url{https://www.kodakprofessional.com/sites/default/files/wysiwyg/pro/resources/f4043_TMax_400.pdf}
\item Kodak Tri-X: \url{https://business.kodakmoments.com/sites/default/files/files/resources/f4017_TriX.pdf}
\end{itemize}





\end{document}
